TrenchR: an R package for transparent environmental and ecological biophysics

### 2022-05-06 ManuscriptTe.Rmd
Minor fixes

#calculate Julian day
date.ct <- strptime(dat$datetime, format="%m/%d/%Y %H:%M", tz="MST")
dat$J= round(as.numeric(julian(date.ct, origin = as.Date("2011-01-01") )))
dat$hour= date.ct$hour

@@

We next illustrate quantifying operative environmental temperatures, which are estimated body temperatures for grasshoppers in the specified environments. We first read and process time series of environmental data. We estimate the zenith angle, the angle between the sun's rays and the vertical, across Julian days and hours.
```{r}
#Calculate zenith
dat$psi= zenith_angle(dat$J,lat=dat$lat, lon=dat$lon, hour=dat$hour)

@@

We then implement the energy budget as follows.
```{r, results = FALSE, warning=FALSE}
TeGates= rep(NA, nrow(dat) )

>>>>

#calculate day of year
date.ct <- strptime(dat$datetime, format="%m/%d/%Y %H:%M", tz="MST")
dat$J= round(as.numeric(julian(date.ct, origin = as.Date("2011-01-01") )))
dat$hour= date.ct$hour

@@

We next illustrate quantifying operative environmental temperatures, which are estimated body temperatures for grasshoppers in the specified environments. We first read and process time series of environmental data. We estimate the zenith angle, the angle between the sun's rays and the vertical, across days of the year and hours.
```{r}
#Calculate zenith
dat$psi= zenith_angle(dat$J,lat=dat$lat, lon=dat$lon, hour=dat$hour)

@@

We then implement the energy budget as follows.
```{r echo=T, results = 'hide', warning=FALSE}
TeGates= rep(NA, nrow(dat) )

### 2022-05-06 Manuscript_full.Rmd
Minor fixes

-   Within the word count (6000-7000 words for Standard Articles, 3000
    words for Applications and Practical Tools)

@@

1.  Much understanding of organimsal responses to climate change and
    variability relies on the assumption that body temperatures are
    equal to temporally averaged air temperatures high above the ground.
    However, most organisms live near the ground and acute exposure to
    solar and thermal radiation and thermal extremes can substantially
    elevate their body temperatures.
2.  We introduce the TrenchR package, which aids in Translating
    Environmental Change into organismal responses. The package includes
    microclimate models to vertically scale weather station data to
    organismal heights. The biophysical modelling tools include both
    general models for heat flows and specific models to predict body
    temperatures for a variety of ectothermic taxa. Additional functions
    model and temporally partition air and soil temperatures and solar
    radiation. We offer utility functions to aid in estimating the
    organismal and environonmental parameters needed for biophysical
    ecology.
3.  TrenchR focuses on simple and modular functions so users can create
    transparent and flexible models for biophysical applications.
4.  The package aims to introduce and enable microclimate and
    biophysical modelling to improve ecological and evolutionary
    forecasting. We further this aim through a series of educational
    modules to introduce the field of biophysical ecology.

@@

The TrenchR package (<http://trenchproject.github.io/TrenchR/>) aims to
promote transparency and reproducibility. We welcome contributions and
corrections from users. Our package is built using the devtools
methodology (<https://github.com/r-lib/devtools>) in Github. Issues and
feature requests can be contributed in Github
(<https://github.com/trenchproject/TrenchR>). The packageis available
via Github and has been submitted to CRAN.

@@

**Specific biophysical functions** provide biophysical models that have
been built for specific organisms based on their physical properties,
behavior, environment, and life history. We currently provide published
biophysical models for lizards, salamanders, butterflies, grasshoppers,
limpets, mussels, and snails. Most models predict operative
environmental temperatures, which are the equilibrium body temperatures
of organisms with specified physical properties in a specific
microclimate and assume no heat exchange via metabolism or evaporation
[@bakken1992measurement].

@@

## Example Application

>>>>

-   Within the word count (6000-7000 words for Standard Articles, 3000-4000
    words for Applications and Practical Tools)

@@

1.  Much understanding of organimsal responses to climate change and
    variability relies on the assumption that body temperatures are
    equal to temporally averaged air temperatures high above the ground.
    However, most organisms live near the ground and acute exposure to
    solar and thermal radiation and thermal extremes can substantially
    elevate their body temperatures.
2.  We introduce the TrenchR package, which aids in Translating
    Environmental Change into organismal responses. The package includes
    microclimate models to vertically scale weather station data to
    organismal heights. The biophysical modelling tools include both
    general models for heat flows and specific models to predict body
    temperatures for a variety of ectothermic taxa. Additional functions
    model and temporally partition air and soil temperatures and solar
    radiation. We offer utility functions to aid in estimating the
    organismal and environonmental parameters needed for biophysical
    ecology.
3.  TrenchR focuses on simple and modular functions so users can create
    transparent and flexible models for biophysical applications.
4.  The package aims to introduce and enable microclimate and
    biophysical modelling to improve ecological and evolutionary
    forecasting. We further this aim through a series of educational
    modules that introduce the field of biophysical ecology.

@@

The TrenchR package (<http://trenchproject.github.io/TrenchR/>) aims to
promote transparency and reproducibility. We welcome contributions and
corrections from users. Our package was built using the devtools
methodology (<https://github.com/r-lib/devtools>) in Github. Issues and
feature requests can be contributed in Github
(<https://github.com/trenchproject/TrenchR>). The package is available
via Github and has been submitted to CRAN.

@@

**Specific biophysical functions** provide biophysical models that have
been built for particular organisms based on their physical properties,
behavior, environment, and life history. We currently provide published
biophysical models for lizards, salamanders, butterflies, grasshoppers,
limpets, mussels, and snails. Most models predict operative
environmental temperatures, which are the equilibrium body temperatures
of organisms with specified physical properties in a specific
microclimate and assume no heat exchange via metabolism or evaporation
[@bakken1992measurement].

@@

## Examples
